### 2015-01-10 paper/Content/01introduction.tex
Minor changes

What brings the actual value to an \emph{Android} device is the installed software, the so called \quotes{Applications} or 
\quotes{Apps}.
Software for \emph{Android} devices can be downloaded and installed via the 
\emph{Play Store}, a software distribution solution, developed by Google and 
pre-installed on every new device, which currently holds over 1.4 million 
applications (as of November 2014) \cite{AppBrainStats}.\\

%\subsection{Application Development}

@@

%\subsection{Challenges in Android Development}

@@

However, not having clear rules regarding the application's architecture entails risks, especially for people that are not expert-level developers. Applications become larger from time to time, and without a reasonable architecture this leads to complex code which is hard to understand and maintain \cite{PassiveMVC, BallOfMud}. Additionally, complex dependencies in the SDK cause messy code and make it hard to exchange specific elements without affecting other elements \cite{PassiveMVC}. That is especially because \emph{Android} tends to lead to classes containing a lot of functionality, which often requires more effort concerning testability and extendability \cite{GangOfFour}. \\

%\subsection{Model-View-Presenter Architecture}

A modern architecture approach is Model-View-Presenter (MVP), a model introduced in 1996 to support event-driven systems \cite{TaligentMVP}. Its main goal is to separate the data logic from the presentation logic, thus creating software which is more flexible in terms of replacing the data model without changing the presentation part or vice versa \cite{PassiveMVC}. Therefor, the software is split into three parts: The \emph{Model} provides the data and means to manipulate it, the \emph{View} part is responsible to display the data, and the \emph{Presenter} contains the business logic that handles user-initiated events and issues the appropriate manipulation of the data \cite{TaligentMVP}.\\

%\subsection{Applying MVP in Android}

Applying MVP to \emph{Android} applications seems promising to build less complex applications with clear code and improved extensibility. However, as the considerations in this paper show, an \emph{Android} implementation of MVP is tricky. Some components of the SDK have several responsibilities and make it hard to define whether they are a \emph{Model}, a \emph{View} or a \emph{Presenter} \cite{PassiveMVC}.

The approach developed in this paper aims to clarify the distinction of these architectural parts. The findings of our research may serve as a guideline on how to apply MVP in \emph{Android} application development.\\

%\subsection{Research Questions}
In the following sections, these questions concerning the application of MVP to \emph{Android} application development are evaluated and answered:

\begin{enumerate}[label={RQ\arabic*:}, leftmargin=0.95cm]
\item What are the inherent characteristics given / recommended by the Android Framework? What is the current state in Android Development in accordance to the inherent framework characteristics?
\item What are the key benefits of the Model-View-Presenter approach?
\item How can the MVP model be applied to the Android App architecture considering framework building blocks?
\end{enumerate}

To answer these questions, a prototypical App is built using the MVP approach.
\todo[inline]{Outlook for each chapter?}

>>>>

What brings the actual value to an \emph{Android} device is the installed software, the so called \quotes{Applications} or 
\quotes{Apps}.
Software for \emph{Android} devices can be downloaded and installed via the 
\emph{Play Store}, a software distribution solution, developed by Google and 
pre-installed on every new device, which currently holds over 1.4 million 
applications (as of November 2014) \cite{AppBrainStats}.

% App Development

@@

% Challenges in Android Development

@@

However, not having clear rules regarding the application's architecture entails risks, especially for people that are not expert-level developers. Applications become larger from time to time, and without a reasonable architecture this leads to complex code which is hard to understand and maintain \cite{PassiveMVC, BallOfMud}. Additionally, complex dependencies in the SDK cause messy code and make it hard to exchange specific elements without affecting other elements \cite{PassiveMVC}. That is especially because \emph{Android} tends to lead to classes containing a lot of functionality, which often requires more effort concerning testability and extendability \cite{GangOfFour}.

% Model-View-Presenter

A modern architecture approach is Model-View-Presenter (MVP), a model introduced in 1996 to support event-driven systems \cite{TaligentMVP}. Its main goal is to separate the Data Management (Model) and User Interface (View-Controller), thus creating software which is more flexible in terms of replacing the \emph{Model} without changing the presentation part or vice versa \cite{PassiveMVC}.
Potel's \emph{Model-View-Presenter} pattern consists of three components: \emph{Model}, \emph{View} and \emph{Presenter}. The \emph{Model} represents the domain model including the business logic. The \emph{View} represents the user interface and interacts with the user and displays the data. The \emph{Presenter} is the mediator between the business logic (Model) and the user interface (View). Its responsibility is to prepare the data (presentation logic), to handle user-initiated events and issues the appropriate manipulation of the data.

% Applying MVP in Android

Applying MVP to \emph{Android} applications seems promising to build less complex applications with clear code and improved extensibility. However, as the considerations in this paper show, an \emph{Android} implementation of MVP is tricky. Some components of the SDK have several responsibilities and make it hard to define whether they are a \emph{Model}, a \emph{View} or a \emph{Presenter}.

The approach developed in this paper aims to clarify the distinction of these architectural parts. The findings of our research may serve as a guideline on how to apply MVP in \emph{Android} application development.

% Research Questions

In the following sections, these questions concerning the application of MVP to \emph{Android} application development are evaluated and answered:

\begin{enumerate}[label={RQ\arabic*:}, leftmargin=0.95cm]
\item What are the inherent characteristics given/recommended by the \emph{Android} framework and what is the current state in \emph{Android} application development in accordance to the inherent framework characteristics?
\item What are the key benefits of the \emph{Model-View-Presenter} approach?
\item How can the MVP approach be applied to the Android application architecture considering framework building blocks?
\end{enumerate}

To answer these questions, a prototypical App is built using the MVP approach.

In the next section we shortly show what core components, provided by the \emph{Android} SDK, need to be considered for our MVP implementation. In addition, we elaborate what types of \emph{Model-View-Presenter} approaches exist and which one we have chosen to build our App prototype. This is followed by an detailed explanation how we applied MVP to the Android SDK and how we built our App prototype. The last section concludes our findings.

### 2015-01-10 paper/Content/02research.tex
Solution chapter
Merges 'images' and 'Images'
App prototype testing
Minor changes

\section{Research}

@@

@@

\textbf{Fragment}: Fragments split more complex parts of an activity's user interface into several smaller sub-components. Fragments are also used to encapsulate reusable parts of the user interface, which for example need to appear in mulitple activities. The official documentation of the \emph{Android} SDK defines a fragment as follows \cite{SDKFragments}:

@@

\subsection{MVP's two flavors: Passive View \& Supervising Controller}

The \emph{Model-View-Presenter} (MVP) pattern was introduced in 1996 by Potel \cite{TaligentMVP} while developing a programming model for multi-tier client/server applications. Potel explains in his paper how to generalize the classical \emph{Model-View-Controller} (MVC) approach in order to build more complex applications. His idea was to decompose the MVC approach into two fundamental concepts: Data Management (Model) and User Interface (View-Controller).

Like in MVC, Potel's \emph{Model-View-Presenter} pattern consists of three components: \emph{Model}, \emph{View} and \emph{Presenter}. The \emph{Model} represents the domain model including the business logic. The \emph{View} represents the user interface and interacts with the user and shows the data from the model in a certain way. The \emph{Presenter} is the mediator between the business logic (Model) and the user interface (View). Its responsibility is to prepare the data (presentation logic) in order to show it in the user interface and vice versa, to determine how user interactions are going to change the model.

\begin{figure}[!t]
\centering
\includegraphics[width=0.5\textwidth]{Images/mvp}
\caption{Supervising Controller and Passive View}
\label{fig:mvp}
\end{figure}

@@

@@

In contrast to the Supervising Controller, the \textbf{Passive View} totally separates the \emph{View} from the \emph{Model}. There is no direct or indirect interaction between them -- see Figure \ref{fig:mvp}. Interactions between \emph{View} and \emph{Model} is done exclusively via the \emph{Presenter}. This approach reduces the \emph{View} to an absolute minimum by using the \emph{Presenter} to handle all user input events, to update the \emph{Model} based on user interactions as well as to channel \emph{Model} changes back to the \emph{View}. As a result, the \emph{View} is made completely passive and is no longer responsible for updating itself from the \emph{Model}.\\
The primary driver for \emph{Passive View} is testing. This concept allows testing to be focused on the \emph{Presenter} without needing any interaction with the chosen UI framework.

\subsection{Approach to solving the problem}

>>>>

\section{MVP applied to Android Apps}

@@

Before we elaborate how we applied the MVP approach to our prototype

In the next section we show in detail how we applied that shit

@@

\textbf{Fragment}: Fragments split more complex parts of an activity's user interface into several smaller sub-components. Fragments are also used to encapsulate reusable parts of the user interface, which for example need to appear in multiple activities. The official documentation of the \emph{Android} SDK defines a fragment as follows \cite{SDKFragments}:

@@

\subsection{MVP: Passive View or Supervising Controller?}

@@

\begin{figure}[!t]
\centering
\includegraphics[width=0.5\textwidth]{mvp}
\caption{Supervising Controller and Passive View}
\label{fig:mvp}
\end{figure}

@@

In contrast to the Supervising Controller, the \textbf{Passive View} totally separates the \emph{View} from the \emph{Model}. There is no direct or indirect interaction between them -- see Figure \ref{fig:mvp}. Interactions between \emph{View} and \emph{Model} is done exclusively via the \emph{Presenter}. This approach reduces the \emph{View} to an absolute minimum by using the \emph{Presenter} to handle all user input events, to update the \emph{Model} based on user interactions as well as to channel \emph{Model} changes back to the \emph{View}. As a result, the \emph{View} is made completely passive and is no longer responsible for updating itself from the \emph{Model}.

The primary driver for \emph{Passive View} is testing. This concept allows testing to be focused on the \emph{Presenter} without needing any interaction with the chosen UI framework.
